---
title: "Econ 10R Week 4"
author: "Luke Zhang"
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Functions in R

There are many functions in R, and we have seen and used many of them like `mean`, `median` and `var`. We have even written our own functions. Today we will be going through more details on functions and their applications.

Most basic functions in R are easy to use - just write out the function's name, for example, `mean`, and follow it up with what we call an argument. The argument is the input in the bracket that follows the function, and it can be raw data, an R object, or results from other functions. For example:

```{r}
die <- 1:6
mean(1:6)
mean(die)
round(mean(die))
```

The function `round` returns the rounded number. It is clear that R resolves the functions of the inner most bracket first, and then move on to the outer ones.

Previously, we talked about a die using the `sample` function. It is quite easy to do in one step:

```{r}
sample(die, 1)
```

Or alternatively

```{r}
sample(1:6, 1)
```

Now, suppose we wish to roll the die twice in a row, we can try the following:

```{r}
sample(die, 2)
```

It seems fine at first, but you will quickly notice that the two numbers will never be the same, and this is wrong - if you roll the dice enough times, you will end up with the same number sooner or later. The reason is that `sample` has set replacement to `FALSE`, so that when the function rolls a number, it does not replace the number rolled - instead it is taken away from the set so that it will not appear again. Clearly we do not want this, so we will now try sampling with replacement.

## Sample with Replacement

The easiest way to add replacement in the example above is to do the following:

```{r}
sample(die, size = 2, replace = TRUE)
```

Adding `replace = True` tells R to put the number back into the set after the first roll, so that it is now possible to draw the same number again in the second row.