---
title: "Econ 10R Week 4"
author: "Luke Zhang"
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Functions in R

There are many functions in R, and we have seen and used many of them like `mean`, `median` and `var`. We have even written our own functions. Today we will be going through more details on functions and their applications.

Most basic functions in R are easy to use - just write out the function's name, for example, `mean`, and follow it up with what we call an argument. The argument is the input in the bracket that follows the function name, and it can be raw data, an R object, or results from other functions. For example:

```{r}

```

The function `round` returns the rounded number. It is clear that R resolves the functions of the inner most bracket first, and then move on to the outer ones.

Previously, we talked about rolling a die using the `sample` function. It is quite easy to do in one step:

```{r}

```

Or alternatively

```{r}

```

Now, suppose we wish to roll the die twice in a row, we can try the following:

```{r}

```

It seems fine at first, but you will quickly notice that the two numbers will never be the same, and this is wrong - if you roll the dice enough times, you will end up with the same number sooner or later. The reason is that `sample` has set replacement to `FALSE`, so that when the function rolls a number, it does not replace the number rolled - instead it is taken away from the set so that it will not appear again. Clearly we do not want this, so we will now try sampling with replacement.

## Sample with Replacement

The easiest way to add replacement in the example above is to do the following:

```{r}

```

Adding `replace = True` tells R to put the number back into the set after the first roll, so that it is now possible to draw the same number again in the second row. Now, suppose I want to have a simple function to simulate rolling two dice, I can try the following:

```{r}

```

The strange thing is that dice returns the same result each time. This is because `dice` only has the results saved from the first roll, and it does not actually re-roll the dice. To do so, we actually need to write our own function for it.

## Writing a function in R

In order to re-roll the dice without re-running the entire code, we can write a simple function:

```{r}

```

Effectively, turning it into a function makes R run `sample` whenever you call for the function `roll`. Notice that you do not always need an argument for the function to work.

We can even write more complicated functions. Recall that in the last question of the midterm, we wanted to filter the observations such that only the ones with the highest index for a particular pollutant remains for each date. Let's try to simplify it here:

```{r}

```

The original code for the pollutant `CO` looks like this:

```{r}

```

And we have to write the same line of code for each pollutant. Instead, let's create a function, call it `clean`:

```{r}

```

Note that `x` here is the argument, and it will a name of the column from the data frame `air`. When we use column names as arguments, R cannot evaluate them until they are first "quoted", and the "unquoted". This is done using the `enquo` function from the `dyplr` package and then using `!!` on the argument itself. Let's replicate the previous result with the function `clean`:

```{r}

```

Now compare the two results:

```{r}

```

They are exactly the same. This means that our function was a success. Now it will much easier to do it for the other pollutants. For example:

```{r}

```

The function `clean` significantly shortened the code we need to write.

So far we have simplified our code by reducing a series of complicated operations in a single function. But there clearly are still space for improvement - we still have to copy and paste things like `colnames` as well as creating new data frames repeatedly. Next, we can try to simplify the code even further by getting rid of repeated calculations over different variables.

# Loops

Loops are functions that repeat a certain operation over and over again, based on your specifications. The most commonly used loop function is the `for` loop. There are several ways to do the loop, and the simplest of them is to loop over the numeric indices.

## Loop Over Numeric Indices

Let me start with creating the following simulated data:

```{r}

```

Now, suppose I wish to get the median of each column. I can do the following:

```{r}

```

This is simple enough. However, imagine if your dataset has many many more columns. Then doing this by hand does not seem so feasible. Instead, we can use the `for` loop. First, let's create a vector where the results will be stored:

```{r}

```

Here we created a new vectors, whose elements are `double` (numbers), and the length is the number of columns in the data frame `df` - each column returns a median, so we need as many space as the number of columns for `df`. Next we can write the loop:

```{r}

```

The index is `i`, and you can change it to `x`, `y` or anything you like really. The command `seq_along()` returns a sequence that is equal to the length of the data frame specified. In this case, `seq_along(df)` is equivalent to `1:4`. However sometimes you do not know the dimension of the data frame, or rather your data frames changes a lot when you update your code, so it is safer to use `seq_along`. The content of the `for` loop is rather easy to understand, but note that we use `[[]]` to insert the index `i` regardless whether it is a vector like `output` or a data frame like `df`, so that the dollar sign is not needed for `df`. The loop runs repeated, and each time a different value for `i` is grabbed from `seq_along(df)`.

## Loop Over Names

Sometimes we need to loop over names rather than indices. We will start with the following simple example:

```{r}

```

We looped over three names and asked R to print them, which is done here. Let's get back to the `air` data. We can use loops to further simplify the solution for the last question of the midterm:

```{r}

```

This is rather complicated so let's look at it step by step. Line #1 grabs the column names of all the pollutants who are columns 6 to 11 in the `air` dataset, and store them in a new vector of character strings called `name`. Line #2 creates a vector of dates only, called `air_index`, and will be useful later on.

Line #3 and #4 are code within the loop. We are looping over the names of the pollutants. Line #3 is the most complicated one. It is wrapped around the `assign` function. Basically, we turn what `clean` does into a data frame using `data.frame`, assign a name to it which is just the pollutant name (hence `paste0("", nm)`). By doing this, we generated six data frames, each with the name of the corresponding pollutant. Notice that we repeatedly used a function called `get`. This is to get rid of the quotation mark around the variable name so that when it runs, it will treat the variables like `CO_24h` rather than `"CO_24h"`, and the latter does not really work.

Line #4 left-joins each newly generated data frame with `air_index`, so that we have all six columns in station codes plus the seventh which has all the dates.

But we have not renamed the columns yet. This is what line #5 does - it renames the columns as the pollutants. The order will not be messed up as we did this with the loop, so that everything is in the order that the `name` vector stores the characters. The function `substr(name, 1, nchar(name) - 4)` gets rid of the `_24h` for us.

Finally, line #6 gets rid of the duplicated dates, so we end up with exactly 730 dates and therefore rows. We did the last question of the midterm in 8 relatively short lines, as opposed to nearly 20 in the original solution. Imagine the additional savings we can make if there are more pollutants to look at!

## For Loops vs. Functions

We can actually incorporate `for` loops into functions. This is what makes R really good. Recall the `df` dataset we created earlier:

```{r}

```

Previously, we used `for` loop to get the median for each column. Now imagine that you wish to do it for many different data frames. We can write a function with the data frame as the argument while incorporating the loop:

```{r}

```

Now we can apply this function to the data frame `df`:

```{r}

```

We can improve this even further. Getting the `median` may not be the only thing we want - we may also want the `mean` and the `var`. How can we make the function flexible enough to cover these additional functions? This is what we can do:

```{r}

```

What did we do? We added another argument to the function called `fun`. Whatever you replace your `fun` with, it will be the function that does the operation. For example:

```{r}

```

The degree of flexibility went up even further. There are many other ways in which functions and loops interact with each other. You can explore on your own.

## Loops with `apply`

There is an alternative way to use loop, and that is by using the `apply` family of functions. Check the R document for details, but basically `apply` works through either columns or rows of a data frame. For example, let's get the median using `apply` instead of `for`:

```{r}

```

If you change `2` to `1` in the function, R will loop over the rows rather than columns:

```{r}

```

Another useful member of the `apply` family is `lapply`. It returns a list rather than. Let's create an interesting list of data frames:

```{r}

```

Now we stored the pollutant data frames in a list. As an exercise, let's loop over the list of data frames and get rid of the duplicated dates:

```{r}

```

Obviously, for what we were trying to achieve in that question, working with `list` may not be optimal. However, this is just to show you how flexible `lapply` can be. If you wish, you can convert the elements in the `air_list` back into data frames.