---
title: "Econ 10R Week 1"
author: "Luke Zhang"
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects in R

Generally speaking, objects are where you store your data. Your data may be in different forms, shapes and dimensions, and assigning them as objects will make it easy for you to reference them in R. Let's take a look.

```{r}
a <- 1
```

We have a piece of information (datum), 1. Now we have assigned this information to an R object, `a`. After this, R knows what `a` contains. Whenever you call for `a`, it will show you what `a` is.

```{r}
a
```

You can even perform certain operations on the object, depending on the nature of the object.

```{r}
a+1
```

This seems simple and useless enough, but when you have a lot more information to store, it becomes much more useful. Consider the case where you want to create a die on your computer. A standard die has six numbers on it, so it will need to store six numbers.

```{r}
die <- 1:6
die
```

You have a lot of liberty in naming objects, but just try to avoid using symbols like `^`, `!`, `$`, `@`, `+`, `-`, `/`, or `*`, as they all have their own uses in a line of code. Also, it is worth remembering that R is case-sensitive, so naming objects with a capital letter will make it a different object entirely.

You can check the names you have used for objects with the following command:

```{r}
ls()
```

Another very useful tool for beginners that is often overlooked is removing objects in the environment. As a beginner, you will make a lot of mistakes when assigning objects. It is then very useful to clear the environment while keeping all the packages you have loaded (we will talk about packages later) so that you can run everything again after correcting for the mistakes in the code. You can use the following command

```{r}
rm(list = ls())
```

If you want to remove one or a few objects only (instead of clearing everything in the environment), you can simply write `rm(object name)`.

You can also do operations on R objects just like numbers. We also saw that you can add 1 to the object `a`. You can do the same thing for the object `die`. However, without further specification, R will assume the operation is pairwise (element-wise) - it will apply the operation to each element in the object. Let us look at an example:

```{r}
a <- 1
die <- 1:6
die+1
```

We can see that each element in `die` increased by 1. If you multiply `die` with `die`, instead of returning a six by six matrix, it will return a vector of the same dimension after element-wise operation:

```{r}
die * die
```

This is the result of simply asking R to multiply the vectors. If you wish to get a matrix, you need more specific command:

```{r}
die %o% die
```

Or to get the dot product

```{r}
die %*% die
```

This course is not about linear algebra, so you do not have to remember all these operations or know what they mean unless you are interested in linear algebra (you should). This is just to tell you the basics in operation on objects in R. Next, we will move on to more specific examples.

## Atomic Vectors

Let's try to create more complicated datasets in R from scratch. Think of an Excel sheet you've seen before. We will be creating one of those from scratch. In Excel, what you have to do is choose a cell and put numbers or texts in them. In R, we code them so that it is easy to track what you have done in each step.

We start with atomic vectors, which are just basic vectors. Recall how we created the `die` vector before. Now let us do it more formally.

```{r}
rm(list = ls())
die <- c(1, 2, 3, 4, 5, 6)
die
```

You can use `c` to create atomic vectors by grouping values together. You can check if an object is a vector by using the `is.vector` command:

```{r}
is.vector(die)
```

Atomic vectors do not need to contain just numbers. They can be a vector of characters as well. For instance:

```{r}
text <- c("32", "Olympic", "Games")
text
```

Notice that the number `32` can also be a character. A simple rule is that when R displays something with quotation mark around them, then they are character strings rather than numbers.

One last thing I want to emphasize is the logical operators. Whenever you use or see `TRUE` and `FALSE` without quotation marks, they act as logical operators. For example:

```{r}
1+1>2
```

So we are told that 1+1 is not greater than 2 by R. What a surprise. These two logical operators are very important later on in the course, when you have to specify whether to enable or disable certain option. We will see in another example later.

## Attributes

Attributes are basically information of the vector that you attach to. Common ones include names (labels), dimensions and classes. For now we focus on names. Names are simply labels of the values in the vector. Take the `die` vector we created earlier as an example. We can first check if there are any names attributes:

```{r}
names(die)
```

Clearly it does not. Now we can assign names to `die`. The way to do it is to create a vector of characters that you wish to name the values of `die`, then assign them to the output of `names(die)`:

```{r}
names(die) <- c("one", "two", "three", "four", "five", "six")
names(die)
```

Now if you check the vector again, it will show both the values and attributes:

```{r}
die
```

To emphasize that names are basically just labels, we can do the following:

```{r}
die+1
```

As you can see, the names correspond to the values only as labels, and are not related to the nature of the values.

## Matrices

Matrices store data in two-dimensional array, same as in matrix in linear algebra. We can create a matrix based on the atomic vector `die` we already have. We then tell R how many rows or columns we want for this matrix:

```{r}
m <- matrix(die, nrow = 2)
m
```

This `m` matrix we created is two (row) by three (column) in dimension, as can be seen through the command `nrow=2`. We can also see that the logic R uses to fill out the matrix is to complete column first. This can be changed by a simple command:

```{r}
m <- matrix(die, nrow = 2, byrow = TRUE)
m
```

Now R fills each row first before moving on to the next. Also, note that we used the logical operator `TRUE` to enable the `byrow` option. Logical operators are most commonly used in this type of scenario.

## Lists

Lists are like atomic vectors, except that they do not group values together - they group objects together. Let us take a look:

```{r}
list1 <- list(die, m, text)
list1
```

It is easy to see that `list` is a great way to group things together in R as there is no restriction on what it can contain - elements in a list do not have to have the same type.

## Data Frame

Data frame is one of, if not the, most powerful way to store data in R. You can think of a data frame as an Excel spreadsheet as they are both two dimensional. A vector will become a column in a data frame. We can create a simple data frame by hand to see how it works:

```{r}
df <- data.frame(face = c("ace", "two", "six"),  
  suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 6))
df
```

Data frames share some similarities with lists. You can create the above data frame in a list as well. Data frame is more restrictive when it comes to its formation, but is also a clearer way to present data. Let us see another example:

```{r}
df2 <- data.frame(Name = c("Bolt", "Powell", "Gatlin"),
                  Record = c(9.58, 9.72, 9.74),
                  Nationality = c("Jamaican", "Jamaican", "American"))
df2
```

We have three observations with three variables. The first observation has `"Bolt"` recorded under the first variable, `9.58` under the second variable and `"Jamaican"` under the third. Everything is tidy and clear.

## Loading and Saving Data

Creating data frames by hand is not practical if you have even just a little bit more data. Not only is it too time consuming, but you will inevitably make typos and mistakes. Moreoever, the data we analyze are usualy from external sources. It is then only natural to learn how to load downloaded data into R, deal with them, and then save them as an output.

We downloaded the `deck.csv` from the HOP textbook. Let's try and load it. First, to simplify your task, just put the file under the same folder as your `Rmd` file. Then we can read the data with the following command:

```{r}
read.csv("deck.csv")
```
Simple enough. However, we cannot do anything with it yet as we only told R to read the dataset, not load it. To do so, we can create a new data frame and load the data into it:

```{r}
deck <- read.csv("deck.csv")
```

Now we have a deck of cards. We can start to manipulate them if we want, but let's just hold off for now. Suppose that you are done with the analysis and want to save your output, we can use the following command:

```{r}
write.csv(deck, file = "cards.csv", row.names = FALSE)
```

Now the data frame `deck` is saved as `cards.csv`. Note that we are simply saving the file into the same path that we are working on, so if you wish to save it in a different place, you can either change the working directory or write out the full path. Also, make sure if you are naming your data right. Do not name your output with the same name as your original, raw data. If you did this while working in the same folder, you might overwrite your raw data, which is something you will never want to do.

The option `row.names = FALSE` prevents R from creating an extra column containing numbers that identify your observations (1 to 52). It is not very useful because most software will just treat it as a column of data.

# Basic Notations in R

So far we have learned several ways of structuring data in R, and we have a fully loaded dataset, `deck`. Now we can play around the dataset a little bit.

## Selecting Values

To select a (set of) specific value from a data frame, we can use the following command:

```{r}
deck[1, 1]
```

The first number in the square bracket is the row you want, and the second number refers to the column. There are several ways to extract an entire row or column:

```{r}
deck[1, ]
deck[1, c(1:3)]
deck[1, c("face", "suit", "value")]
```

All three commands return the same thing. If you wish to ignore a particular column or row, you can simply drop it from the selection:

```{r}
deck[1, c(1:2)]
deck[1, c(1, 3)]
```

You can also do the same thing with a negative sign:

```{r}
deck[1, c(1:2)]
deck[1, -3]
```

When a minus sign is added, it excludes rather than includes the corresponding row or column. If we want to drop the king of spades only, we can do:

```{r}
deck[-1, ]
```

## Deal Cards with Functions

Now we know how to extract a certain card, we can write a function to deal a card. We can simply as R to return the one on the top of the deck:

```{r}
deal <- function(x){
  x[1, ]
}
deal(deck)
```

However, if you run this function over and over, it will keep returning the king of spades to you. This is because neither does r know that king of spades should be taken away aftering being selected, nor do we shuffle the deck afterwards. It is more difficult to do the former, so let's try the latter for now.

## Shuffle Cards with Functions

Before we jump straight to shuffling the deck, let us try to use what we learned so far to reorganize the deck the way we want. Suppose I want to move the second card from top (second row) to the first and keep everything else unchanged:

```{r}
deck[c(2, 1, 3:52), ]
```

This is simple: I wrote down specifically the order of the rows (row 2 first, row 1 second, and then 3 to 52), and left the column blank so every column is included. This is no where near random shuffling. However, this gives us a hint: as long as we can randomly order the 52 rows, we can assign the new order to the deck and have it shuffled. To do so, we use the following command:

```{r}
random <- sample(1:52, size = 52)
random
```

And let's apply this new order to the deck:

```{r}
deck[random, ]
```
Now we can shuffle and create a new deck using a `shuffle` function:

```{r}
shuffle <- function(x){
  random <- sample(1:52, size = 52)
  x[random, ]
}
deck2 <- shuffle(deck)
deck2
```

Now we can deal and get a different card:

```{r}
deal(deck2)
```

We can even combine the two steps and skip the creationg of a new data frame (deck2):

```{r}
deal(shuffle(deck))

deal(shuffle(deck))

deal(shuffle(deck))
```

You can see that we have different result every time. True shuffling it is.

## Dollar Sign

A quick note on how to reference a particular column in a data frame or a list. Usually in economics, we will assign labels to variables. We can use the label and dollar signs as reference. For example, if I wish to extract the first column of the deck (the column that has the face of the cards), we can simply write:

```{r}
deck$face
```

We can incorporate this when we do calculations. For example:

```{r}
mean(deck$value)
```

For lists, we can do the same:

```{r}
names(list1) <- c("One", "Two", "Three")
list1$One
```

# Modifying Values in R

Most of the time, the data you load into R are not perfect like the one we just used. You will encounter typos, non-numeric values in the place where numbers should  be, missing values and extreme values. Furthermore, as you may load data in many different formats (e.g., `csv`, `xlsx` and `dta`), each may have different way of treating missing observations and therefore causing R to react to them differently. Hence it is imperative that we learn how to modify (usually on a small scale) values in R.

## Changing values

Let us use the `die` vector as examples. Suppose we want to change the first value in `die` to 1000, we can use the following command:

```{r}
die[1] <- 1000
die
```

You can also replace a group of values:

```{r}
die[c(1, 3, 5)] <- c(1000, 3000, 5000)
die
```

And doing certain operations on a group of values:

```{r}
die[c(2, 4, 6)] <- die[c(2, 4, 6)] + 1
die
```

You can add observations by directly adding values. For example, the `die` vector used to have only 6 values, but now we can add a seventh one:

```{r}
die[7] <- 0
die
```

You can also add a new variable to a data frame in the same way, remember to use the dollar sign:

```{r}
deck2$newvar <- c(1)
head(deck2)
```

If you wish to remove a particular value or variable, we can use the `NULL` symbol:

```{r}
deck2$newvar <- NULL
head(deck2)
```
Now let's try something slightly more complex. We know that aces in poker are generally considered to be of the highest value, so their values should not just be 1. Instead, we should assign them a value that is greater than 13, say 14. In the original deck, it is not very difficult to do as you know exactly where each ace appears:

```{r}
deck$value[c(13, 26, 39, 52)]
deck$value[c(13, 26, 39, 52)] <- c(14, 14, 14, 14)
deck$value[c(13, 26, 39, 52)]
```

However in the shuffled deck, it is not as easy, since we do not really know where the aces are. As a result, we need to replace values **conditional** on their faces being aces. To do this, we can make use of the `==` syntax, which is R's way to test equality. Do not confuse this with `<-`, which is the operator for assigning objects.

```{r}
deck2$value[deck2$face == "ace"] <- 14
deck2[deck2$face == "ace", ]
```
Once you understand the logic of these operations, it is very easy to make up your own "conditions". There are other symbols like `&`, `!` and `|` that can be very useful when you make up conditions, and we will surely be using them later on in the course.

## Dealing with Missing Values

As a data scientist, dealing with missing values is our bread and butter. As an economist, it is even more so as economic data are never perfect. Usually the first step you complete after loading the data is to check for missing data and, if appropriate, remove them. Let me first generate some missing values:

```{r}
random2 <- sample(1:52, size = 5)
random2
deck2$value[c(random2)] <- c(NA, NA, NA, NA, NA)
deck2
```

Now we can see that `deck2` have 5 `NA` (missing value) under the variable `value`. Now suppose that this deck is not for poker, just a random dataset that you have. Also assume that the missing values are random (it indeed is, as I generated the missing values randomly), we can then drop them in order to continue our analysis. We can check what values are missing by using the `is.na` option:

```{r}
is.na(deck2)
```

Clearly, we have a few missing values under the variable `value` as expected. R told us this by showing `TRUE` whenever there is a missing value. Now let's remove the missing values for good. Remember that we will be dropping **observations** that have missing values, not just dropping the missing values. Therefore, cards without a `value` will be dropped even if they still have their `face` and `suit` non-missing. Again, you do not want to mess up the original data frame as you can make mistakes, so let us create a new data frame which contains the non-missing observations.

```{r}
deck3 <- na.omit(deck2)
is.na(deck3)
dim(deck3)
```

The new data frame `deck3` has observations removed if they have a missing value.

With these basic tools of value modification in hand, we can move on to deal with more complicated data. Before that, let us take a step back and talk about packages and libraries in R.

# Installing Packages and Loading Libraries

So far we have used only very basic commands in R. In order to use more advanced and convenient commands, we sometimes need to install packages. Packages are collections of functions, data and documentation in R that greatly extend R's basic capabilities. Here let us look at a very useful package called `tidyverse`. It is more like a collection of packages, and makes cleaning data a lot easier.

Before we can use anything from `tidyverse`, we will need to install it. This is because some functions of `tidyverse` is not included in the basic R. To install `tidyverse` (and other packages in general), we can use the following command:

```{r eval=FALSE, include=TRUE}
install.packages("tidyverse", type = "binary")
```

THe installation might take a while, and you may need to respond to some prompts. You may see some errors in the process. But as long as you see the package downloaded (it should end with a line `The downloaded source packages are in`), then do not worry too much about it. We will deal with it later if we encounter problems.

Now that we have the package installed, you can exclude it from your codes, as you do not need (or want) to install it everytime you run the codes - it just needs to be installed once.

After successful installation, we need to load the relevant functions in order to use them. To do so we can do the following:

```{r}
library(tidyverse)
```

Generally you only need to load the library once as well. However loading library does not take a long time so you can keep it in the code if you want. For the conflicts you might see, this is because some functions in `tidyverse` share the same name as basic functions in R. If you want to use the base functions of R, you need to write `stats::filter()` instead of `filter()`, as the function with the same name from `tidyverse` is masking the base function.

We can install other packages and load other libraries in a similar way. Having learned the codes, let us move back to data.

# Data Transformation and Visualization

To learn more about how to clean and visualize data, we need to use a more complicated data set. Let us load the following library:

```{r eval=FALSE, include=TRUE}
install.packages("nycflights13")
```

```{r}
library(nycflights13)
```



This loads a dataset on flights departing New York City in 2013. Let us load the data as a data frame:

```{r}
nycflights13::flights
```

We can see that this dataset is a lot more complex than `deck`. It has 336,776 observations, each with 19 variables. This dataset contains variables of many different natures. Some of them are integers, denoted by `int`, some are character strings, denoted by `chr`, et cetera. Now let us mess around with the data a little bit.

## Filter

We already learned how to select certain observations from a dataset. However, it is usually the case that this is more than one way to do the same thing. Here we can make sure of the `filter` command to do the same thing in a simpler way. Suppose we wish to select all the flights on the day of Jan 1st:

```{r}
filter(flights, month == 1, day == 1)
```

Now only flights from Jan 1st appears and the sample size shrinks to 842, a good start. Comparing to using `[]` to subset the data, `filter` is a little bit more intuitive and can do things more easily. Right now we only have the data shown to us but not stored in objects, so to do that, we can assign the selected data to a new data frame:

```{r}
jan1 <- filter(flights, month == 1, day == 1)
```

You can easily click it in the environment and view it. There are other comparison operators you can use to filter the data. For example, if you wish to include flights in November or December, we can write:

```{r}
filter(flights, month == 11 | month == 12)
```

If you wish to find flights that had a delay (both arrival and departure) of no more than two hours, you can write the following:

```{r}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

Looks like the sample size did not shrink a lot, but no more than two hours is a pretty slack condition. Alternatively, we can achieve the same thing with the following command:

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
```

The operator `!` means "not something". Remember that `arr_delay > 120 | dep_delay > 120` means arrival delay of more than two hours **or** departure delays of more than two hours. Adding `!` excludes all that, so logically it means the same thing as `arr_delay <= 120, dep_delay <= 120`. To sum up, apart from standard notations like `>`, `<`, `>=`, `<=`, `==`, operators like `,` (and), `!` (not) and `|` (or) can also be very useful when setting filtering conditions.

## Arrange

You can arrange rows by variables. For example, if you wish to sort the flights by year-month-day, we can do the following:

```{r}
arrange(flights, year, month, day)
```
In our data this does not change much as this is the original way the dataset was arranged. R defaults to sorting by ascending order, but you can switch to descending order if you wish by doing this:

```{r}
arrange(flights, desc(dep_delay))
```

By default, missing values (`NA`) are always sorted at the end. It is always worth remembering that none of the above arrangement is really saved unless you assign the output into a new data frame.

## Select

When you want to select just a few variables out of potential hundreds, you can use the `select`:

```{r}
select(flights, year, month, day)
```

We now have 336,776 observations but only three variables. The sample size shrank a lot. The `select` command is very useful for truncating datasets if you just want to get rid of certain variables. There are other options you can use. For example, you can pick all the variables between year and day (inclusive) by doing the following:

```{r}
select(flights, year:day)
```

You can also exclude all the variables between year and day (inclusive):

```{r}
select(flights, -(year:day))
```

There are other conditions that you can set. For example, using `starts_with("abc")` will select all variables that start with `abc`, `ends_with("xyz")` and `contains("efg")` work similarly. There are even more options that you can easily google up.

## Mutate

You will often come across situations where you wish to create a new variable that is a function of existing variables. This is where we can use the `mutate` option. Let us take a look. First we will create a dataset with fewer variables so that it is easy to see everything.

```{r}
flights_short <- select(flights, 
                        year:day,
                        ends_with("delay"),
                        distance,
                        air_time)
```

Now suppose we want to know the time gained in air (delay time minus arrival time) and the average speed of the flight:

```{r}
mutate(flights_short,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60)
```

Two new variables are created and amended to the end of the data frame. You can create other variables in similar fashion, and there are also many operations you can do on existing variables that can generate new variables (e.g., leads and lags). Obviously we cannot cover everything here, but learning how to add variables in an important first step.

## Summarize

Sometimes the unit of observation is too small or too detailed for your purpose (yes, that can happen). In those cases, You will want to collapse a set of rows into a single row. For example, it is very tedious to analyze every single flight in the data frame, so maybe we will want to look at the data day by day. What we can do is to take the average of the flights of a day. For example, to get the average delay per date:

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, count = n(), delay = mean(dep_delay, na.rm = TRUE))
```

First, we used `group_by` to change the unit of analysis to each date (think about what would happen if I leave out "year" and "month"), so the `by_day` dataset has the same observation but with a different unit of analysis. Then I collapsed the data by date and got the average departure delay. The command `summarise` does the collapsing for us. Also note that we applied `na.rm = TRUE` so that we ignore all missing values. Try on your own what would happen if we do not use this option.

I included a new variable called `count = n()` here. This gives you the number of observations that you are doing your operation (in our case, `mean()`) on. So for example, on January 1st, we are averaging the delay for 842 flights. This number is not very useful to use right now, but in cases where the number of observations in a group is small enough to affect the outcome, knowing the number would be useful. We will come across examples in the future.

If you wish to ungroup a dataset, simply apply:

```{r}
ungroup(by_day)
```

## Group Mutates and Filters

You can also use the `group_by` function to mutate the data frame. For example, suppose we wish to find the worst offenders of arrival delays by day, we can do the following:

```{r}
flights_short <- group_by(flights_short, year, month, day)
flights_short <- filter(flights_short, rank(desc(arr_delay)) <= 10)
flights_short
```

Now what do we have? We picked out the top 10 worst delays for each day. We can also find all groups bigger or less than a threshold:

```{r}
popular_dest <- filter(group_by(flights, dest), n() > 365)
popular_dest
```

What is this? We first changed the unit of analysis to each destination. Then we kept all the flights which have a destination that has on average more than 1 flight per day from New York and deleted the rest. Note that I incorporated the `group_by` function under `filter`, saving an extra step compared to the previous code where we were finding the worst delays by day.